---
title: "Assignment1_Sep14"
author: "Tejaswini Yeruva"
date: "2022-09-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Packages

Packages that were installed for the below:

```{r eval=TRUE}

require(tidyverse)

library(tidyverse)

```
### Dataset Source
Dataset has been sourced from Data world.

```{r Importing DataSet}
travel_data <- read.csv("https://query.data.world/s/hsiafhxaakgh3yvwfxxk3dvkyrjhex", header=TRUE, stringsAsFactors=FALSE);

print(travel_data)
```
### Statistics description
Below is the summary of the dataset.
```{r Descriptive}
summary(travel_data)
```
### Transforming Variable in the Dataset
Dataset has been sorted with Duration of Stay

```{r Descriptive Statistics}
travel_data_new <-travel_data[order(travel_data$Duration.of.stay),]
print(travel_data_new)
```

### Box Plot and Scatter Plot
Boxplot has been made with Duration of Stay and Scatter plot has been made with Duration of Stay and Total Transport costs.


```{r Plots}
boxplot(travel_data$Duration.of.stay, main='Duration of Stay',
        col.axis = "darkgreen", col.lab = "darkgreen")

plot(x = travel_data$Duration.of.stay, y = travel_data$Total.transport.costs, xlab = "Duration.of.stay",
     ylab = "Total.transport.costs", main = "Cost VS Stay")
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.